---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

f1dataR serves as a tool to get neatly organized Formula 1 data into your R environment. Here we will go over the basic functions to understand how the package works.

The most sought-after aspect of F1 data is telemetry data. 
Let's get Leclerc's fastest lap from the first race of 2022:
```{r setup}
library(f1dataR)

get_driver_telemetry(2022, 1, driver = 'LEC', fastest_only = T)
```


now let's use ggplot2 to visualize some of the data we have

```{r}
library(tidyverse)

lec <- get_driver_telemetry(2022, 1, driver = 'LEC', fastest_only = T) %>% 
  head(300)%>% 
  mutate(index = 1:300)

ggplot(lec, aes(index, Throttle))+
  geom_line()+
  theme_minimal()
```

What if we get more drivers involved. Let's also get the Quaifiying data from Hamilton

```{r}
ham <- get_driver_telemetry(2022, 1, 'Q', driver = 'HAM', fastest_only = T) %>% 
  head(300) %>% 
  mutate(index = 1:300)

per <- get_driver_telemetry(2022, 1, driver = 'PER', fastest_only = T) %>% 
  head(300)%>% 
  mutate(index = 1:300)

data <- bind_rows(lec,ham,per)

ggplot(data, aes(index, Throttle, color = driverId))+
  geom_line()+
  theme_minimal()

```

We can also use one of the plotting functions included in the package.

```{r}
plot_fastest(2022, 1, 'R', 'PER')
```

