---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

#Use a local temporary cache and set back to user's info when finished vignette build
withr::local_file(file.path(getwd(), "vignettes","vig_build_tmp"))
if(dir.exists(file.path(getwd(), "vignettes","vig_build_tmp"))){
  unlink(file.path(getwd(), "vignettes","vig_build_tmp"), recursive = T, force = T)
}
dir.create(file.path(getwd(), "vignettes","vig_build_tmp"), recursive = TRUE)
withr::local_options(f1dataR.cache = file.path(getwd(), "vignettes","vig_build_tmp"))
```

f1dataR serves as a tool to get neatly organized Formula 1 data into your R environment. Here we will go over the basic functions to understand how the package works.



The most sought-after aspect of F1 data is telemetry data. 
Let's get Leclerc's fastest lap from the first race of 2022:
```{r setup}
library(f1dataR)

load_driver_telemtery(2022, 1, driver = 'LEC', fastest_only = T)
```


now let's use ggplot2 to visualize some of the data we have

```{r}
library(tidyverse)

lec <- load_driver_telemtery(2022, 1, driver = 'LEC', fastest_only = T) %>% 
  head(300)%>% 
  mutate(index = 1:300,
         driver_id = "LEC")

ggplot(lec, aes(index, throttle))+
  geom_line()+
  theme_minimal()
```

What if we get more drivers involved. Let's also get the Qualifying data from Hamilton

```{r}
ham <- load_driver_telemtery(2022, 1, 'Q', driver = 'HAM', fastest_only = T) %>% 
  head(300) %>% 
  mutate(index = 1:300,
         driverId = "HAM")

per <- load_driver_telemtery(2022, 1, driver = 'PER', fastest_only = T) %>% 
  head(300)%>% 
  mutate(index = 1:300,
         driver_id = "PER")

data <- bind_rows(lec,ham,per)

ggplot(data, aes(index, throttle, color = driver_id))+
  geom_line()+
  theme_minimal()

```

Now lets visualize speed and use the f1 dark theme for ggplot included in the package

```{r}
ggplot(data, aes(index, speed, color = driver_id))+
  geom_line()+
  theme_dark_f1()+
  theme(
    axis.title = element_text(),
    axis.line = element_line(color = 'white'),
  )+
  labs(
    title = "Speed Comparison for Bahrain 2022"
  )
```


We can also use one of the plotting functions included in the package.

```{r}
plot_fastest(2022, 1, 'R', 'PER')
```

