---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

#Use a local temporary cache and set back to user's info when finished vignette build
withr::local_file(file.path(getwd(), "vignettes","vig_build_tmp"))
if(dir.exists(file.path(getwd(), "vignettes","vig_build_tmp"))){
  unlink(file.path(getwd(), "vignettes","vig_build_tmp"), recursive = T, force = T)
}
dir.create(file.path(getwd(), "vignettes","vig_build_tmp"), recursive = TRUE)
withr::local_options(f1dataR.cache = file.path(getwd(), "vignettes","vig_build_tmp"))
```

f1dataR serves as a tool to get neatly organized Formula 1 data into your R environment. Here we will go over the basic functions to understand how the package works.



The most sought-after aspect of F1 data is telemetry data. 
Let's get Leclerc's fastest lap from the first race of 2022:
```{r telem}
library(f1dataR)

load_driver_telemetry(2022, 1, driver = 'LEC', fastest_only = T)
```

Now let's use ggplot2 to visualize some of the data we have

```{r simple_plot}
library(dplyr)
library(ggplot2)

lec <- load_driver_telemetry(2022, 1, driver = 'LEC', fastest_only = T) %>% 
  head(300)

ggplot(lec, aes(distance, throttle))+
  geom_line()+
  theme_minimal()
```

What if we get more drivers involved. Let's also get the Qualifying data from Hamilton and Perez

```{r triple_plot}
ham <- load_driver_telemetry(2022, 1, 'Q', driver = 'HAM', fastest_only = T) %>% 
  head(300)

per <- load_driver_telemetry(2022, 1, driver = 'PER', fastest_only = T) %>% 
  head(300)

data <- bind_rows(lec,ham,per)

ggplot(data, aes(distance, throttle, color = driver_code))+
  geom_line()+
  theme_minimal()

```

Now lets visualize speed and use the f1 dark theme for ggplot included in the package

```{r use_theme}
ggplot(data, aes(distance, speed, color = driver_code))+
  geom_line()+
  theme_dark_f1()+
  theme(
    axis.title = element_text(),
    axis.line = element_line(color = 'white'),
  )+
  labs(
    title = "Speed Comparison for Bahrain 2022"
  )
```


We can also use one of the plotting functions included in the package.

```{r use_builtin}
plot_fastest(2022, 1, 'R', 'PER')
```



Now let's look at a more complete analysis. We want to visualize how lap time change over time (tyre age) for Pérez with every compound used in the Spanish GP.

```{r}
laps <- load_session_laps(2023, 'Spain') %>% 
  filter(driver == 'PER') %>% 
  group_by(compound) %>% 
  filter(tyre_life != 1 & tyre_life != max(tyre_life)) %>% # Remove in and out laps
  ungroup()
ggplot(laps, aes(tyre_life, lap_time, color = compound))+
  geom_line()+
  geom_point()+
  theme_dark_f1(axis_marks = T)+
  labs(
    color = "Tyre Compound",
    y = "Lap Time (Seconds)",
    x = "Tyre Life (Laps)"
  )+
  scale_color_manual(
    values = c('white', 'yellow', 'red')
  )+
  scale_y_continuous(breaks = seq(75,85, 1))+
  scale_x_continuous(breaks = seq(2,26, 4))

```

